Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-30 manuscripts/reply/reply-main.tex
Finishing draft of reply ms overhaul.

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method was very sensitive to prior assumptions and often
incorrectly supported shared divergences across taxa that diverged randomly
over broad time periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
% The papers of \citet{Oaks2012} and \citet{Hickerson2013} are in agreement
% on the fundamental methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{itemize}
    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
        rejection algorithm implemented in \msb is unable to adequately sample
        the space of the models within reasonable computational time.
        They propose the method incorrectly supports models with fewer
        divergence-time parameters because they are better sampled.
    \item \citet{Oaks2012} suggest the cause is more fundamental; the broad
        uniform priors simply lead to very small marginal likelihoods of models
        with many divergence-time parameters.
        Models with more divergence-time parameters integrate over \emph{much}
        greater parameter space with low probability of producing the data, yet
        relatively high prior density, resulting in small marginal likelihoods.
\end{itemize}
In the former case, the problem is estimation error; the true (correct)
posterior supports models with many divergence-time parameters given data
generated under temporally random divergences, but we are unable to accurately
estimate the posterior.
In the latter case, estimation error is not the issue, because the true
posterior supports simultaneous divergences given data from randomly diverged
taxa; this is the correct (albeit biogeographically misleading) answer given
the prior assumptions about divergence times imposed by broad uniform priors.
Such posterior support for simultaneous divergence, even if correct from the
perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.

While these phenomena are not mutually exclusive, it is important to
distinguish between them in order to determine how best to improve our ability
to estimate shared divergence histories.
If the cause is estimation error, then the model is sound and we need to
improve our numerical estimation procedures.
For example, \citet{Hickerson2013} proposed a model-averaging approach that
uses narrow, empirically informed uniform priors in the hope that with less
parameter space to sample, the rejection algorithm will produce better
estimates of the posterior.
If the cause is marginal likelihoods, we need to correct the model; improving
the estimation machinery will only increase our ability to estimate
biogeographically misleading posteriors.
Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without overly inhibiting the marginal likelihoods of models
with more divergence-time parameters, which increases the method's robustness
and power to detect temporal variation in divergences.

Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice and evaluate the empirical Bayesian
model-averaging approach of \citet{Hickerson2013} as a potential solution to
the biases of \msb revealed by \citet{Oaks2012}.
In their re-analysis of the dataset of \citet{Oaks2012}, \citet{Hickerson2013}
made an error by mixing different units of time, which invalidates the results
presented in their response (see Supporting Information for details).
We avoid this error, but find their approach will often incorrectly support
(1) clustered divergence models and
(2) models that exclude from consideration the true values of the models'
parameters.
% Our results show that it is difficult to choose a uniformly distributed prior
% on divergence times that is broad enough to confidently contain the true
% parameter values while being narrow enough to avoid spurious support for models
% with less parameter space.
We also explore the distinct predictions made by the alternative causes of the
bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to incorrectly infer shared
divergences is the uniform priors on divergence times leading to small marginal
likelihoods of models with more divergence-time parameters.
We also show that sampling error is not a major contributer to the problem.
This is not surprising given the rich statistical literature showing that
marginal likelihoods are very sensitive to the priors used in Bayesian model
selection
\citep[e.g.,][]{Jeffreys1939,Lindley1957}.
As predicted by this fundamental principle of Bayesian model selection,
alternative prior assumptions can improve robustness of the method and greatly
mitigate misleading support for models of shared divergences
\citet{Oaks2014dpp}.

\section{The potential implications of empirical Bayesian model choice}
\change{
\citet{Oaks2012} found \msb incorrectly supports models of temporally
clustered divergences even when using a prior on divergence times that was
informed by the data being analyzed.
The main argument of \citet{Hickerson2013} is that the priors used by
\citet{Oaks2012} were not informative enough.
}
%
They suggest a very narrow, highly informed uniform prior on divergence times
is necessary to avoid the method's preference for models with few
divergence-time parameters.
Such an empirical Bayesian approach to model selection raises some theoretical
and practical concerns, some of which were discussed by \citet{Oaks2012} (see
the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}); we expand on this here.

@@

Thus, empirical Bayes methods have an uncertain theoretical basis and
do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
empirical Bayesian estimates of the posterior are often too narrow, off-center,
and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
This is not to say that empirical Bayesian approaches are not useful.
Empirical Bayes is a well-studied branch of Bayesian statistics that has given
rise to many methods for obtaining parameter estimates that often
exhibit favorable frequentist properties.
% Furthermore, many post-hoc correction methods have been developed for
% estimating confidence-intervals from empirical Bayes estimates of
% distributions that often exhibit well-behaved frequentist coverage
% probabilities
\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The marginal likelihood of the model only appears as a normalizing constant in
the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
under which the data are probable and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior density is no longer a normalizing constant, rather it is how the data
inform the posterior probability of the model.
Because the prior probability distributions placed on the model's parameters
have a strong affect on the integrated, or ``average'', likelihood of a model,
Bayesian model choice tends to be much more sensitive to priors than parameter
estimation \citep{Jeffreys1939,Lindley1957}.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the posterior
probability of a model, is not a function of \myTheta{}, because the parameters
are integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice) will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is no such justification for model selection, because unlike
model parameters, the posterior probabilities of candidate models often have no
clear true values.
% There is no such justification for empirical Bayesian model choice, because
% there are no true values for the model probabilities being estimated.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

@@

\subsection{Practical concerns about empirically informed uniform priors for
    Bayesian model choice}
In addition to the theoretical concerns discussed above, there are practical
problems with using narrow, empirically informed, uniform priors.
% for a method that has already been shown to be biased toward models with less
% parameter space.
The results of Hickerson et al.'s (\citeyear{Hickerson2013}) reanalysis of the
Philippine dataset strongly favored models with the narrowest, empirically
informed prior on divergence times, and thus their model-averaged posterior
estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
\citet{Hickerson2013}).
This is concerning, because the narrowest \divt{} prior used by
\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
divergence times for at least some of the Philippine taxa.
%, a major problem when using uniform priors.
\citet{Hickerson2013} set this prior to match the 95\% highest posterior
density (HPD) interval for the mean divergence time estimated under one of the
priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
Given this interval estimate is for the \emph{mean} divergence time across all
22 taxa, it may be inappropriate to set this as the limit on the prior, because
some of the taxon pairs are expected to have diverged at times older than the
upper limit.
Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
of the mean divergence time under the other two priors explored by
\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
see Table~6 of \citet{Oaks2012}).

\change{
The strong preference for the narrowest prior on divergence times suggests the
approach of \citet{Hickerson2013} is biased toward models with less parameter
space and, as a consequence, will estimate model-averaged posteriors dominated
by models that exclude true values of the model's parameters.
}
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
erroneous model yielding a posterior that is incorrect (i.e., the
model-averaged posterior will be dominated by an incorrect model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

For our re-analyses of the Philippines dataset we followed the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid their error of mixing units of time (see SI for details).
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence-time parameters:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics after standardizing the
statistics using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ model with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors exclude
most, if not all, of the divergence times of the Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
strongly prefers the prior model with the narrowest distribution on divergence
times across all three of our analyses, even when this model is incorrect and
excludes the true divergence times of the Philippine taxa
(Table~\ref{tabModelChoiceEmpirical}).
However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior-predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior-predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these incorrect priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

% Given that the results of \citet{Hickerson2013} strongly prefer the models with
% the narrowest prior on divergence times, it seems quite likely that their
% model-averaged results are dominated by models that exclude true divergence
% times, making their results difficult to interpret.

@@

To better quantify the propensity of Hickerson et al.'s
(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
datasets in which the divergence times for the 22-population pairs are drawn
randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
Exp(2)$).
All other parameters were identically distributed as the $M_1$--$M_5$ models
(Table~\ref{tabModelChoiceEmpirical}).
We then repeated the analysis described above using $1\e6$ random samples from
each of the prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$, retaining 1000
posterior samples for each of the 1000 simulated datasets.

For each simulation replicate, we estimated the Bayes factor in favor
of excluding the truth as the ratio of the posterior to prior odds of
excluding the true value of at least one parameter.
Whenever the Bayes factor preferred a model excluding the truth we counted the
number of the 22 true divergence times that were excluded by the preferred
model.
Our results show that the model-averaging approach of \citet{Hickerson2013}
favors a model that excludes the true values of parameters in 97\% of the
replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
true divergence times (Figure~\ref{figExclusionSimTau}).
We also used the mode estimate of the preferred model for each replicate to
estimate the number of true parameter values excluded, which produced nearly
identical results.
Importantly, the posterior probability of excluding at least one true parameter
value is very high in most replicates
(Figure~\ref{figExclusionSimProb}).
Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
of the replicates (87\% with GLM-regression adjustment) strongly support the
exclusion of true values (Figure~\ref{figExclusionSimProb}).

The results of the above empirical and simulation analyses clearly demonstrate
the risk of using narrow, empirically guided uniform priors in a Bayesian
model-averaging framework.
The consequence of this approach is obtaining a model-averaged posterior
estimate that is heavily weighted toward incorrect models that exclude true
values of model parameters.
% \subsection{Additional thoughts on empirical priors in Bayesian model choice}
This is not a general critique of Bayesian model averaging.
Rather, model averaging can provide an elegant way of incorporating
model uncertainty in Bayesian inference.
However, when averaging over models with narrow and broad uniform priors on a
parameter with a likelihood density that is not expected to be uniformly
distributed, the posterior can be dominated by models that exclude from
consideration the true values of parameters due to the larger marginal
likelihoods of the models that integrate over less space with high prior
weight yet low likelihood.

\change{
When using uniformly distributed priors, the alternative to capturing prior
uncertainty is to risk excluding the true values one seeks to estimate.
Fortunately, more flexible continuous distributions that are better suited as
priors for the positive real-valued parameters of the \msb model have been
shown to greatly reduce spurious support for clustered divergence models while
allowing prior uncertainty to be accommodated
\citep{Oaks2014dpp}.
}

% Given the theoretical and practical issues with empirical Bayes approaches to
% Bayesian model choice, it is clear why one should use caution before overly
% criticizing an investigator's choice of priors \emph{after} having seen the
% resulting posterior.
% As discussed by \citet{Oaks2012}, prior to analyzing the data, there was a
% large amount of uncertainty regarding the divergence times of the 22 population
% pairs under study.
% Two of these pairs represent distinct species, and the taxonomy of many groups
% in the Philippines has repeatedly been shown to mask deeply divergent lineages
% \citep{RafeDiesmosAlcala2008,Linkem2010,Siler2010,Welton2010,Siler2011HerpMonographs,
% Siler2011,Siler2012,RafeStuart2012,LinkemBrown2013,Rafe2013AREES,Siler2014kikuchii}.

\section{Assessing the power of the model-averaging approach of
    \citet{Hickerson2013}}
While our results above clearly demonstrate the risks inherent to the empirical
Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
such risk if the approach does indeed increase power to detect temporal
variation in divergences and thus decrease bias toward clustered divergence
models.
We assess this possibility using simulations.
Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
the 22 population pairs randomly drawn from a uniform distribution, $U(0,
\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
\globalcoalunit generations.
All other parameters were identically distributed as the prior models.
As above, we generated a prior sample of $5\e6$ total samples from the five
prior models $M_1$, $M_2$, $M_3$, $M_4$, and $M_5$
(Table~\ref{tabModelChoiceEmpirical}).
For each simulated dataset, we approximated the posterior
by retaining 1000 samples from the prior with the smallest Euclidean distance
from the true summary statistics, as described above.
In total, we analyzed 6000 replicate datasets, retaining 1000 model-averaged
posterior samples for each of them.

@@

% The method is most likely to infer the extreme case of a single divergence
% event shared across all 22 taxa when populations diverged randomly over the
% past \globalcoalunit generations.
% When divergences are random over the past $8\globalpopsize$ generations, the
% most likely inference is only two divergence events, and a single
% divergence is still estimated in more than 10\% of the replicates.
% It is interesting to note that as \divt{max} increases, but before the
% estimates are finally pulled away from $\numt{} = 1$, the distribution of
% \numt{} estimates closely mirror the ``U-shaped'' prior on divergence models
% used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
% (\citeyear{Oaks2012}) Figure 5B).
% This suggests the U-shaped prior coupled with the small marginal likelihoods
% of models with many divergence parameters is a major cause of the method's
% bias toward clustered divergence models; this is consistent with
% \citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.

>>>>

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method was very sensitive to prior assumptions and often
supports shared divergences across taxa that diverged randomly over broad time
periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{itemize}
    \item \citet{Hickerson2013} suggest that under broad uniform priors, the
        rejection algorithm implemented in \msb is unable to adequately sample
        the space of the models within reasonable computational time, which
        leads to support for models with fewer divergence-time parameters
        because they are better sampled.
    \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
        priors on divergence times simply lead to very small marginal
        likelihoods of models with many divergence-time parameters. These
        models are ``averaged'' over much greater parameter space with large
        prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
\end{itemize}
In the former case, the problem is estimation error; the true (correct)
posterior supports models with many divergence-time parameters given data
generated under temporally random divergences, but we are unable to accurately
estimate the posterior.
In the latter case, estimation error is not the issue, because the true
posterior supports simultaneous divergences given data from randomly diverged
taxa.
I.e., this is the correct (albeit biogeographically misleading) answer given
the prior assumption that divergence times are broadly and uniformly
distributed.
Such posterior support for simultaneous divergence, even if correct from the
perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.

While these phenomena are not mutually exclusive, it is important to
distinguish between them in order to determine how best to improve our ability
to estimate shared divergence histories.
If the cause is estimation error, then the model is sound and we need to
improve our numerical estimation procedures.
For example, Markov chain or sequential Monte Carlo algorithms might improve
the efficiency of sampling from the posterior.
Rather than alter the sampling algorithm, \citet{Hickerson2013} tried a
model-averaging approach that uses narrow, empirically informed uniform priors
in the hope that with less parameter space to sample, the rejection algorithm
will produce better estimates of the posterior.
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice and evaluate the approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their analyses, \citet{Hickerson2013} made an error by mixing different
units of time, which invalidates the results presented in their response (see
Supporting Information for details).
We correct this error, but still find their approach will often support
(1) clustered divergence models when divergences are random, and
(2) models that exclude from consideration the true values of the
parameters.
The behavior of their approach also suggests that estimation error is not the
primary cause of the issue.

If support for models with few divergence events is caused by their larger
marginal likelihoods, we need to correct the model; improving the estimation
machinery will only increase our ability to estimate true posteriors that are
difficult to interpret.
Accordingly, \citet{Oaks2014dpp} has introduced a method that uses more
flexible probability distributions to accommodate prior uncertainty in
divergence times without overly inhibiting the marginal likelihoods of models
with more divergence-time parameters.
This greatly increases the method's robustness and power to detect temporal
variation in divergences \citep{Oaks2014dpp}.
This is not surprising given the rich statistical literature showing that
marginal likelihoods are very sensitive to the priors used in Bayesian model
selection
\citep[e.g.,][]{Jeffreys1939,Lindley1957}.

We also explore the distinct predictions made by the alternative causes of the
bias proposed by \citet{Oaks2012} and \citet{Hickerson2013}.
We show the primary cause of the method's tendency to support models of shared
divergences is the large marginal likelihoods of these models under the prior
assumption of uniformly distributed divergence times.
We also show that sampling error is not a major contributer to the problem.

\section{The potential implications of empirical Bayesian model choice}
\citet{Oaks2012} found \msb spuriously supports models of temporally
clustered divergences even when using a prior on divergence times that was
informed by the data being analyzed.
The main argument of \citet{Hickerson2013} is that the priors used by
\citet{Oaks2012} were not informative enough.
They suggest a very narrow, highly informed uniform prior on divergence times
is necessary to avoid the method's preference for models with few
divergence-time parameters.
Such an empirical Bayesian approach to model selection raises some theoretical
and practical concerns, some of which were discussed by \citet{Oaks2012} (see
the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}); we expand on this here.

@@

Thus, empirical Bayes methods have an uncertain theoretical basis and
do not yield a valid posterior distribution from Bayes' rule \citep[e.g.,
empirical Bayesian estimates of the posterior are often too narrow, off-center,
and incorrectly shaped;][]{Morris1983,Laird1987,Carlin1990,Efron2013}.
This is not to say that empirical Bayesian approaches are not useful.
Empirical Bayes is a well-studied branch of Bayesian statistics that has given
rise to many methods for obtaining parameter estimates that often
exhibit favorable frequentist properties.
\citep{Morris1983,Laird1987,Laird1989, Carlin1990,Hwang2009}.

@@

In Equation \ref{eq:bayesrule}, we see that the posterior density of any state
$\myTheta{}$ of the model, is the prior density updated by the probability of
the data given $\myTheta{}$ (the likelihood of $\myTheta{}$).
The marginal likelihood of the model only appears as a normalizing constant in
the denominator.
Thus, as long as the prior distribution contains the values of $\myTheta{}$
under which the data are probable and the data are strongly informative
relative to the prior, the values of the parameters that maximize the posterior
distribution will be relatively robust to prior choice, even if the posterior
is technically incorrect due to using the data to inform the priors.
However, if we look at Equation~\ref{eq:bayesmodelchoice}, we see that in
Bayesian model choice it is now the \emph{marginal} likelihood of a model that
updates the prior to yield the model's posterior probability.
The integral over the entire parameter space of the likelihood weighted by the
prior density is no longer a normalizing constant, rather it is how the data
inform the posterior probability of the model.
Because the prior probability distributions placed on the model's parameters
have a strong affect on the integrated, or ``average'', likelihood of a model,
Bayesian model choice tends to be much more sensitive to priors than parameter
estimation \citep{Jeffreys1939,Lindley1957}.
Another important difference of Bayesian model choice illustrated by
Equation~\ref{eq:bayesmodelchoice} is that the value of interest, the posterior
probability of a model, is not a function of \myTheta{} because the parameters
are integrated out of the marginal likelihoods of the candidate models.
Thus, unlike parameter estimates, the estimated posterior probability of a
model is a single value (rather than a distribution) lacking a measure of
posterior uncertainty.

The justification for an empirical Bayesian approach to parameter estimation is
that giving the data more weight relative to the prior (i.e., using the data
twice) will often shift the peak of the estimated distribution nearer to the
true value(s) of the model's parameter(s).
However, there is no such justification for model selection, because unlike
model parameters, the posterior probabilities of candidate models often have no
clear true values.
Model posterior probabilities are inherently measures of our belief in the
models after our prior beliefs are updated by the data being analyzed.
This complicates the meaning of model posterior probabilities when Bayes' rule
is violated by informing priors with the same data to be analyzed.
By using the data twice, we fail to account for prior uncertainty and mislead
our posterior beliefs in the models being compared; we will be over confident
in some models and under confident in others.

@@

\subsection{Practical concerns about empirically informed uniform priors for
    Bayesian model choice}
In addition to the theoretical concerns discussed above, there are practical
problems with using narrow, empirically informed, uniform priors.
The results of Hickerson et al.'s (\citeyear{Hickerson2013}) reanalysis of the
Philippine dataset strongly favored models with the narrowest, empirically
informed prior on divergence times, and thus their model-averaged posterior
estimates are dominated by models $M_1$ and $M_2$ (see Table 1 of
\citet{Hickerson2013}).
This is concerning, because the narrowest \divt{} prior used by
\citet{Hickerson2013} ($\divt{} \sim U(0,0.1)$) likely excludes the true
divergence times for at least some of the Philippine taxa.
\citet{Hickerson2013} set this prior to match the 95\% highest posterior
density (HPD) interval for the mean divergence time estimated under one of the
priors used by \citet{Oaks2012} (see Tables 2 and 3 of \citet{Oaks2012}).
Given this interval estimate is for the \emph{mean} divergence time across all
22 taxa, it may be inappropriate to set this as the limit on the prior, because
some of the taxon pairs are expected to have diverged at times older than the
upper limit.
Furthermore, this prior is \emph{excluded} from the 95\% HPD interval estimates
of the mean divergence time under the other two priors explored by
\citet{Oaks2012} (under these priors the 95\% HPD is approximately 0.3--0.6;
see Table~6 of \citet{Oaks2012}).

The strong preference for the narrowest prior on divergence times suggests the
approach of \citet{Hickerson2013} is biased toward models with less parameter
space and, as a consequence, will estimate model-averaged posteriors dominated
by models that exclude true values of the parameters.
We explored this possibility in two ways.
First, we re-analyzed the Philippines dataset using the model-averaging
approach of \citet{Hickerson2013}, but set one of the prior models with a
uniform prior on divergence times that is unrealistically narrow and almost
certainly excludes most, if not all, of the true divergence times of the 22
taxon pairs.
If small likelihoods of large models cause the method to prefer models with
less parameter space, we expect \msb will preferentially sample from this
erroneous model yielding a posterior that is misleading (i.e., the
model-averaged posterior will be dominated by a model that excludes
the truth).
Second, we generated simulated datasets for which the divergence times are
drawn from an exponential distribution and applied the approach of
\citet{Hickerson2013} to each of them to see how often the method excludes the
truth.

@@

For our re-analyses of the Philippines dataset we followed the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid their error of mixing units of time (see SI for details).
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence times:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics after standardizing the
statistics using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ model with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors exclude
most, if not all, of the divergence times of the Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
strongly prefers the prior model with the narrowest distribution on divergence
times across all three of our analyses, even when this model excludes the true
divergence times of the Philippine taxa
(Table~\ref{tabModelChoiceEmpirical}).
However, \citet{Hickerson2013} vetted the priors used in their model-averaging
approach via ``graphical checks,'' in which the summary statistics from 1000
random samples of each prior model are plotted along the first two orthogonal
axes of a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior-predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ models are problematic, we performed these graphical checks on our
prior models.
Unfortunately, these prior-predictive checks provide no warning that these
priors are too narrow (Figure~S\ref{figPCA}).
Rather, the graphs suggest these invalid priors are ``better fit''
(Figure~S\ref{figPCA}A--C) than the valid priors similar to those used by
\citet{Oaks2012} (Figure~S\ref{figPCA}D--F).

@@

To better quantify the propensity of Hickerson et al.'s
(\citeyear{Hickerson2013}) approach to exclude the truth, we simulated 1000
datasets in which the divergence times for the 22-population pairs are drawn
randomly from an exponential distribution with a mean of 0.5 ($\divt{} \sim
Exp(2)$).
All other parameters were identically distributed as the $M_1$--$M_5$ models
(Table~\ref{tabModelChoiceEmpirical}).
We then repeated the model-averaging  analysis described above, retaining 1000
posterior samples for each of the 1000 simulated datasets.
For each simulation replicate, we estimated the Bayes factor in favor
of excluding the truth as the ratio of the posterior to prior odds of
excluding the true value of at least one parameter.
Whenever the Bayes factor preferred a model excluding the truth we counted the
number of the 22 true divergence times that were excluded by the preferred
model.

Our results show that the model-averaging approach of \citet{Hickerson2013}
favors a model that excludes the true values of parameters in 97\% of the
replicates (90\% with GLM-regression adjustment), excluding up to 21 of the 22
true divergence times (Figure~\ref{figExclusionSimTau}).
Importantly, the posterior probability of excluding at least one true parameter
value is very high in most replicates
(Figure~\ref{figExclusionSimProb}).
Using a Bayes factor of greater than 10 as a criterion for strong support, 66\%
of the replicates (87\% with GLM-regression adjustment) strongly support the
exclusion of true values (Figure~\ref{figExclusionSimProb}).

The results of the above empirical and simulation analyses clearly demonstrate
the risk of using narrow, empirically guided uniform priors in a Bayesian
model-averaging framework.
The consequence of this approach is obtaining a model-averaged posterior
estimate that is heavily weighted toward models that exclude true
values of the parameters.
This is not a general critique of Bayesian model averaging.
Rather, model averaging can provide an elegant way of incorporating
model uncertainty in Bayesian inference.
However, when averaging over models with narrow and broad uniform priors on a
parameter with a likelihood density that is not expected to be uniformly
distributed, the posterior can be dominated by models that exclude from
consideration the true values of parameters due to the larger marginal
likelihoods of the models that integrate over less space with high prior
weight yet low likelihood.

When using uniformly distributed priors, the alternative to capturing prior
uncertainty is to risk excluding the true values one seeks to estimate.
Fortunately, more flexible continuous distributions that are better suited as
priors for the positive real-valued parameters of the \msb model have been
shown to greatly reduce spurious support for clustered divergence models while
allowing prior uncertainty to be accommodated
\citep{Oaks2014dpp}.

\section{Assessing the power of the model-averaging approach of
    \citet{Hickerson2013}}
While our results above clearly demonstrate the risks inherent to the empirical
Bayesian model-choice approach used by \citet{Hickerson2013}, one could justify
such risk if the approach does indeed increase power to detect temporal
variation in divergences.
We assess this possibility using simulations.
Following \citet{Oaks2012}, we simulated 1000 datasets with \divt{} for each of
the 22 population pairs randomly drawn from a uniform distribution, $U(0,
\divt{max})$, where \divt{max} was set to: 0.2, 0.4, 0.6, 0.8, 1.0, and 2.0, in
\globalcoalunit generations.
All other parameters were identically distributed as the prior models.
As above, we generated $5\e6$ samples from the prior models in
Table~\ref{tabModelChoiceEmpirical}.
For each of the 6000 simulated datasets, we approximated the posterior
by retaining 1000 samples from the prior.

@@

Also, the results of our power analyses further demonstrate the propensity of
Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
parameter values.
Across all but one of the \divt{max} we simulated, the method favors a model
that excludes the truth in a large proportion of replicates, and across many of
the \divt{max} the preferred model will exclude a large proportion of the true
divergence times (Figure~\ref{figPowerExclusion4}A--D \&
S\ref{figPowerExclusion6}A--F).
Importantly, the posterior probability of excluding at least one true
divergence value is also quite high across many of the \divt{max}
(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
